---
title: "Lab3"
author: "Mike Loranty"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This week we are going to continue learning how to manipulate and analyze our data. To do this we'll need to use a few new *packages*. Packages in R are groups of functions, typically designed for a specific purpose, that come with detailed documentation on how to use them, and often sampe data sets. For example there are packages designed to work with geospatial data, perform specific statistical tests, and many other things. Today we are going to use two packages called lubridate and dplyr that are useful for working with dates and manipulating data, respectively.

In RStudio you can either install packages from using commands in the console, or using the Tools menu in the toolbar. We will use commands to do this: 

```{r, eval = F}
install.packages(c("dplyr", "lubridate"))
```

This downloads our packages, but now we need to load them into memory for our R session. We do this using the library command. In general it is good practice to load any packages you need in the first lines of your scripts. 

```{r, collapse = T,message = F}
library(dplyr)
library(lubridate)
```

Now we have these packages loaded in memory in R. We'll continue today using data from Siberia archived at the Arctic Data Center: https://arcticdata.io/catalog/view/doi:10.18739/A2S46H57S 

Let's start again by looking at air temperature. 
```{r, collapse = T}
# read in our air temperature data
t_air <- read.csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A3e962856-7ece-4ff8-870a-82511c31c974",header = T)
```

Today we are going to use several functions in the *lubridate* package to work with the dates in our data. First we need to combine all of the date and time attributes into a single character vector. In addition we need to separate out the hours and minutues. We can use the `paste` function to combine the year, doy, hour, and minute. Remember R will operate on entire vectors. Why do we use the `floor` function, what does it do? 

```{r, collapse = T}
# create a character vector with date and time information
d <- paste(t_air$year,t_air$doy,floor(t_air$hour),(t_air$hour-floor(t_air$hour))*60, sep = "_")

# have a look at our new vector
head(d)
```

Now we can use the `as_datetime` function to convert our character vector into a special kind of date object. Remember we use the `$` operator to add new variables (columns) to a dataframe.
```{r, collapse = T}
# convert our charcter vector to a date/time obejct and assign the output to a new variable in our dataframe. The format argument is where we tell R how the date is arranged.
t_air$dt <- as_datetime(d,format = "%Y_%j_%H_%M")

# have a look at our new date and time column
head(t_air$dt)

# use the month function to create a month variable
t_air$month <- month(t_air$dt)

```

Now we can use the functions from the dplyr package to aggregate our date to monthly values. In addition to these new functions, we'll be using a new operator `%>%` to send the output from one command right into the next, this is called *piping*.

```{r, collapse = T}
# create a new object called t_month that is identical to our t_air dataframe and pipe it to the next line
t_month <- t_air %>%

# group the data by year, month, site, and sensor height, and pipe to the next line  
  group_by(year,month,site,sensorZ) %>%
  
# summarize data - calculate the mean and standard deviation for number of observations for each month
  summarize(t_mean = mean(Temp,na.rm=T),
            t_sd = sd(Temp,na.rm=T),
            n_obs <- n())

# check out the new data frame
head(t_month)
```

